
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 13: Writing a function - Conditional Statement"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment13.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

#### NOTICE: Question 7 and 8 are optional. 

-------

1. Write a function to calculate the age of a person.  Give an example to test your function. 

- Input: a number, x, (year born)

- Output: Print out "You are r age". Where r is the age of the person, i.e. 2020 - x. 
```{r}
birthday <- function(x)
{r = 2020 - x
print(paste0("You are ",r," years old!"))}

birthday(1999)
```

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#29)

-------

2. Write the following function. Give an example to test your function. 

- Input: a number

- Output: print out: "You input an even number!" if the number is event, or "You input an odd number!" otherwise. 


```{r}
isItEven <- function(x){
  if(x%%2 == 0)
  {
    print(paste0(x, " is an even number!"))
  }
  else{
    print(paste0(x, " is an odd number!"))
  }
}

isItEven(10)
isItEven(11)
```


Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#33)

-------

3. Write the following function.  Give an example to test your function. 

- Input: a numeric vector 

- Output:  

  - if the input vector has missing values: return the input vector with missing values replaced by mean
  
  - if the input vector has no missing value: return the same input vector
```{r}
nullvector <- function(x){
  if(is.null(x)){
    mean_of_x <- mean(table(x))
    library(tidyr)
    x <- replace_na(x, mean_of_x)
  }
  return(x)
}

p <- c(3, 3, 4, 5, 6, 7, 7)

nullvector(p)

```
  
  
Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#37)

-------

4. Write the following function.  Give an example to test your function. 

- Input: a vector x 

- Output:  The vector x where the missing values replaced by the mean (if x is numeric) or the mode (if x is non-numeric).  If x does not have missing value, return the same vector x.  

```{r}
nullvector <- function(x){
  if(is.numeric(x)){
    if(is.null(x)){
      mean_of_x <- mean(table(x))
      library(tidyr)
      x <- replace_na(x, mean_of_x)
    }
    else if(!is.numeric(x)){
      mode_of_x <- names(sort(-table(x)))[1]
      library(tidyr)
    x <- replace_na(x, mode_of_x) 
    }
    }
  return(x)
}

```


Hint: Combine the function in Question 3 and [this function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#37)

-------

5. Write the following function. Give examples to test your function. 

- Input: A data frame of two variables x and y

- Output: 

  - A boxplot of x by y if x is numeric and y is non-numeric
 
  - A boxplot of y by x if y is numeric and x is non-numeric
 
  - print out 'This function cannot visualize your data' otherwise

Hint: 

- [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#41)
 
- You can refer to this slide to plot a boxplot: https://bryantstats.github.io/math421/slides/6_viz.html#36
 
-------
 
6. Combine the function in Question 5 and the function in this example ([Link](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#41)) to have a function that can plot all possible cases (2 categorical, 2 numeric, 1 numeric and 1 categorical). 

-------

#### NOTICE: Question 7 and 8 are optional. 

-------

7. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
  
  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the accuracy of the decision model `rpart` where the training data is train_percent. 

-------

8. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
  
  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the plot of variable important by random forest trained by caret. 



