
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 12: Predictive Modeling - Part 3"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment12.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Install the package `mlbench` and use the follows to import the data

```{r}
library(mlbench)
library(gapminder)
library(gganimate)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(knitr)
library(caret)
data(PimaIndiansDiabetes)
df <- PimaIndiansDiabetes
```

- Set seed to be 2020.
```{r}
set.seed(2020)
```

- The target variable is `diabetes`
- Partition the data into 80% training and 20% testing.  

```{r}
splitIndex <- createDataPartition(df$diabetes, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```


-------

2. Use cross-validation of 30 folds to tune random forest (method='rf').  What is the `mtry` value that produces the greatest accuracy?

```{r}
tuneGrid = expand.grid(mtry = 2:8)


trControl = trainControl(method = "cv",
                         number = 30)


forest_cv <- train(diabetes~., data=df_train, 
                                method = "rf", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
```


```{r}
#Plot the results
#It looks like an Mtry value of 2 holds the highest accuracy
plot(forest_cv)
```

 
-------

3. Use cross-validation with of 30 folds to tune random forest (method='ranger').  What are the parameters that produce the greatest accuracy?

```{r}
trControl = trainControl(method = "cv",
                         number = 30)

tuneGrid = expand.grid(mtry = 2:4,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:10))


ranger_cv <- train(diabetes~., data=df_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
```

```{r}
plot(ranger_cv)
```


-------

4. Go to https://topepo.github.io/caret/available-models.html and pick a classification model.  Tune the classification model using cross-validation of 30 folds. 

5. Compare the three models in question 2, 3, and 4 to select the final model.  Evaluate the accuracy of the final model on the test data. 
