
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 8: Visualization - Making a bar race"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment8.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Install two packages `gganimate` and `gifski` then restart Rstudio. Use the WHO's dataset at [this link](https://covid19.who.int/WHO-COVID-19-global-data.csv). Make a top-10 bar race by months between countries on the number of deaths by Covid 19.  

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(gganimate)
library(gifski)
df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')

df$month <- month(df$Date_reported, label = TRUE)

```


```{r}
d1 <- df %>% 
  filter(Country %in% c('Ecuador','United States of America'))%>% 
  group_by(month, Country) %>% 
  summarise(mean = mean(Cumulative_deaths)) 

d2 <- d1 %>% 
  group_by(month) %>% 
  mutate(rank=rank(-mean))
```


```{r}
p1 <- d2 %>% 
  ggplot(aes(y=factor(rank), 
             x=mean, group=Country, 
             fill=Country)) + 
  geom_col()+
  transition_states(month)+
  labs(x = 'Total Dead', 
       title = 'Month: {closest_state}')

animate(p1, nframes=100, fps=15)
```


2. Make another bar race using that dataset.  

```{r}
BR2 <- df %>% 
  filter(Country %in% c('Italy','United States of America'))%>% 
  group_by(month, Country) %>% 
  summarise(mean = mean(Cumulative_cases)) 

Bar2 <- BR2 %>% 
  group_by(month) %>% 
  mutate(rank=rank(-mean))
```


```{r}
animate2 <- Bar2 %>% 
  ggplot(aes(y=factor(rank), 
             x=mean, group=Country, 
             fill=Country)) + 
  geom_col()+
  transition_states(month)+
  labs(x = 'Total Cases Over Time', 
       title = 'Month: {closest_state}')

animate(animate2, nframes=100, fps=15)
```


3. Make a bar race using a dataset of your own interest. You may use the dataset that we use in class (https://covidtracking.com/data/download/all-states-history.csv) but you should make a different bar racev from ones in the slides. 

```{r}
df2 <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')

df2$month <- month(df2$date, label = TRUE)
```


```{r}
d1 <- df2 %>% 
  filter(state %in% c('MA','FL'))%>% 
  group_by(month, state) %>% 
  summarise(mean = mean(hospitalized)) 

d2 <- d1 %>% 
  group_by(month) %>% 
  mutate(rank=rank(-mean))
```

```{r}
p1 <- d2 %>% 
  ggplot(aes(y=factor(rank), 
             x=mean, group=state, 
             fill=state)) + 
  geom_col()+
  transition_states(month)+
  labs(x = 'Total Hospitalized', 
       title = 'Month: {closest_state}')

animate(p1, nframes=100, fps=15)
```


