
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7: Visualization - Some aspects of ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment7.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.   
 
-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**Questions**

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis. 

-------

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')

df$day <- wday(df$date)
df$day_type <- case_when(df$day < 6 ~ 'weekday',
                         TRUE~'weekend')

```


1. Make a plot using `face_wrap` by one categorical variable (`face_wrap(~variable1)`)

```{r}
df %>% filter (state=='MA' |state=='RI') %>% 
  ggplot(mapping=aes(x=death, y=hospitalized))+ geom_point()+ facet_wrap(~state) + labs(x='Number of deaths',
                          y="Number of Hospitalizations",
                          title="Deaths vs Hospitalizations",
                          caption="Displays the Deaths versus hospitalizations")
```


2. Make a plot using `face_wrap` by two categorical variables (`face_wrap(variable1~variable2)`)

```{r}
df %>% filter (state=='MA'|state=='RI') %>% 
  ggplot(mapping=aes(x=death, y=hospitalized, color=state))+ geom_point()+ facet_wrap(state~day_type)+
  labs(x='Number of deaths',
      y="Number of Hospitalizations",
      color="States",
      title="Deaths vs Hospitalizations for MA and RI by day type",
      caption="Displays the Deaths versus hospitalizations on a facet wrap")
```


3. Make a plot using `position ='jitter'`. 

```{r}
df %>% filter (state=='MA', day=='7' | day=='3') %>% 
  ggplot()+ geom_point(mapping=aes(x=death, y=hospitalized, color=day_type), position = 'jitter')+
  labs(x='Number of deaths',
      y="Number of Hospitalizations",
      color="Day Type",
      title="Deaths vs Hospitalizations for Sunday and Wed",
      caption="Displays the Deaths versus hospitalizations for two days") 
```


4. Make a plot using `position ='dodge'`.

```{r}
df %>% filter (state=='MA') %>% 
  ggplot()+ geom_bar(mapping=aes(x=dataQualityGrade, fill=day_type), position = 'dodge') +
  labs(x = "Data Quality",
    y='Count of Quality Grade',
      color="Day Type",
      title="Amount of Data Quality Grades",
      caption="Bar chart for Data Quality")
```


5. Make a plot using `position ='fill'`. 

```{r}
df %>% filter (state=='MA') %>% 
  ggplot()+ geom_bar(mapping=aes(x=dataQualityGrade, fill=day_type), position = 'fill') +
  labs(x = "Data Quality",
    y='Count of Quality Grade',
      color="Day Type",
      title="Amount of Data Quality Grades",
      caption="Bar chart for Data Quality")
```


6. Make a plot using `geom_col`.

```{r}
df %>% group_by(dataQualityGrade) %>% summarise(mean=mean(death)) %>%
  ggplot()+ geom_col(mapping=aes(x=dataQualityGrade, y = mean)) +
  labs(x = "Data Quality",
    y='Average deaths',
      color="Day Type",
      title="Amount of Data Quality Grades",
      caption="Column chart for Data Quality")
```


7. Change the theme of one of your plots. 

```{r}
df %>% filter (state=='MA') %>% 
  ggplot()+ geom_bar(mapping=aes(x=dataQualityGrade, fill=day_type), position = 'dodge') +
  labs(x = "Data Quality",
    y='Count of Quality Grade',
      color="Day Type",
      title="Amount of Data Quality Grades",
      caption="Bar chart for Data Quality")+ theme_dark()
```


8. Save a plot and put a link of the plot to your Github Webpage.  To link a plot `abc.png` you just need to insert the following on the README.md

```{r}
df %>% filter (state=='MA') %>% 
  ggplot()+ geom_bar(mapping=aes(x=dataQualityGrade, fill=day_type), position = 'dodge') +
  labs(x = "Data Quality",
    y='Count of Quality Grade',
      color="Day Type",
      title="Amount of Data Quality Grades",
      caption="Bar chart for Data Quality") + ggsave('abc.png')+ggsave('abc.pdf')
```


```{r, eval=FALSE}
[Link to the plot](abc.png)
```





