***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard under Assignment 3. 

-------

Problems
-------
 \

1. Use `read.csv` to import the Covid19 data from WHO: [link](https://covid19.who.int/WHO-COVID-19-global-data.csv)
```{r}
covidDF <- read.csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
View(covidDF)

```

2. How many columns and rows the data have? How many categorical variables the data have?
```{r}
#There are 42,025 rows and 8 columns in this data frame, of these 8 variables 4 of them are categorical
dim(covidDF)
str(covidDF)
```

3. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
#There is 187 total missing values, all of these missing values come from the country_code field
sum(is.na(covidDF))
colSums(is.na(covidDF))
```

4. Correct the types of the variables that are mismatched type.
```{r}
#Fixing data types 
str(covidDF)
class(covidDF$ï..Date_reported)
view(df)

library(lubridate)
covidDF$ï..Date_reported <- ymd(covidDF$ï..Date_reported)
class(covidDF$ï..Date_reported)
```

5. Find the daily average number of cases globally.
```{r}
#On average there was 114568.9 new cases globally on a daily basis
mean(covidDF$New_cases)
mean(by(covidDF$New_cases, covidDF$ï..Date_reported, sum))
```

6. Find the maximum cases a day.  
```{r}
#There was a max of 309532 cases in a day
max(covidDF$New_cases)
max(by(covidDF$New_cases, covidDF$ï..Date_reported, sum))
```

7. Which country at which date has the maximum number of cases a day? *Hint:  use the `which.max` function to find the index of the row that contains the maximum number of cases. Then from the row index, find that country and the date. * 
```{r}
#On September 12th it says India had 97,000 new cases
covidDF[which.max(covidDF$New_cases), ]
```


8.  How many WHO's region there are in the data?
```{r}
#There are 7 WHO regions
length(unique(covidDF$WHO_region))
```

9.  Find the daily average cases in EUROPE. Find the daily average cases by regions.
```{r}
#There is a daily average of 20917 cases a day in Europe, and the following average cases by region
euro <- covidDF[covidDF$WHO_region=="EURO", ]
mean(by(euro$New_cases, euro$ï..Date_reported, sum))
```

10. Find the daily average number of cases in the US in the last 5 days. 

11. Which day of the week has the most number of cases in South East Asia (SEARO region)?
