
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Fall 2020 - Math 421 - Midterm"
---

-------

# Instruction

The midterm has two components: the Rmarkdown notebook (html) and the presentation.  The notebook and the presentation file are due Nov 26. We will do the presentation in class 

**The notebook:** The notebook should be created using `rmarkdown` (like other assignments). The notebook should have a title. 


**The Presentation:** Present your results in 5-10 minutes. To make the presentation using Rmarkdown, do the follows: 

    - In Rstudio -> File -> New File -> R markdown
    
    - In the left panel, click to Presentation -> Click OK
    
    - Now you have an Rmarkdown that can be knitted to be a html presentation 

**What to present**:

  - Present Part 2 - Visualization
  
  - Present Question Question 4, 5 and 6 in Part 3.  
  
  - Present any errors/challenges you run into and how you fix/overcome them. 

**Data:**  

The data for the mid-term project is the Rhode Island Department of Health Hospital Discharge Data.  Each row of the data presents a patient. 

Link: https://drive.google.com/open?id=15QNBf6YYKocK2nNIfpKDer58kQnCPNZJ 

-------

## I. Data Wranggling

1. Download the data file `hdd0318cy.sas7bdat`.  

2. Use `read_sas` in library `haven` to read the data. 
```{r}
getwd()
setwd('/Users/student/Documents/GitHub/NoahCederholm.github.io')
library(gapminder)
library(gganimate)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(knitr)
library(rpart)
library(caret)

finaldf <- read_csv('midterm.csv')
```

    
3. Filter the data to have only patients of the year 2018 (`yod==2018`)
    
4. Select to work with only following variables: 

```{r, eval=FALSE}
#This filter was done and then wrote into the csv file above
#"yod", "payfix","pay_ub92","age",                "sex","raceethn","provider","moa", "yoa","mod","admtype", "asource" , "preopday" ,"los", "service" , "icu","ccu",    "dispub92", "payer"  ,"drg","trandb", "randbg","randbs","orr", "anes","seq", "lab","dtest", "ther","blood","phar", "other","patcon","bwght","total","tot" , "ecodub92","b_wt","pt_state","diag_adm","ancilar" , "campus","er_fee","er_chrg","er_mode","obs_chrg", "obs_hour","psycchrg","nicu_day"
```
 

*Notice*:  You may want to save the current data to your computer for easy access later.  To save the data file use `write_csv(df, 'midterm.csv')`, for example.  

5. What are variables that have missing values?
```{r}
#Variables missing include: payfix, preopday, obs_hour, nicu_day.
colSums(is.na(finaldf))

```

 
6. Remove all variables with missing values
```{r}
finaldf$payfix <- NULL
finaldf$nicu_day <- NULL
finaldf$preopday <- NULL
finaldf$obs_hour <- NULL
```


 
7. Refer to the data description in the file `HDD2015-18cy6-20-19.docx`, which variable recording the month of admission?, which variable recording the month of discharge?
```{r}
#MOA is the variable that notes the month of admission where as MOD is the month of discharge.
```


8. Which month admitted the most number of patients? Which month admitted the most number of male patients?
```{r}
#Month 10 (October) has the most admitted patients with 11408.
finaldf %>% group_by(moa) %>% count()
```


9. Which month has the most number of teenage female patients?
```{r}
#March (The third month) has the most teen female patients with 240
finaldf$AgeGroup <- case_when(
    finaldf$age == 0 ~ 'newborn',
    finaldf$age < 12 ~ 'child',
    finaldf$age < 20 ~ 'teen',
    finaldf$age < 35 ~ 'YoungAdult',
    finaldf$age < 60 ~ 'MiddleAge',
    finaldf$age > 60 ~ 'Old')
finaldf %>% filter(sex== '2', AgeGroup == 'teen') %>% group_by(moa) %>% count()
```


10. Which provider has the most number of female patients in October? 
```{r}
#Provider 7205 has the most female patients in october with 1517
finaldf %>% filter(sex== '2', moa == '10') %>% group_by(provider) %>% count()

```

11. Is female patients older than male patients, on average? 
```{r}
#Male patients are older on average with an age of 51.5 and female at 50.9
finaldf %>% group_by(sex) %>% summarise(mean(age))
```


12. Calculate the average age of patients by months. Which month has the oldest patients on average age?
```{r}
#January has the oldest average age at 51.8
finaldf %>% group_by(moa) %>% summarise(mean(age))
```


13. What is the name of the provider that has the highest total charge?
```{r}
# Provider 7205 (Rhode Island Hospital) has the highest total charge at $775,984,880
finaldf$trandb <- as.numeric(finaldf$trandb)
finaldf %>% group_by(provider) %>% summarise(sum(trandb))
```

14. What is the name of the provider that has the least total charge for teenage male on average?
```{r}
#Provider 7214 (Woman and Infants) have the least teen male patients charge on average with 3500
finaldf %>% filter(sex == 1, AgeGroup == 'teen') %>% group_by(provider) %>% summarise(mean(trandb))
```

15. Calculate the length of stays by races.  Which race has the longest length of stays on average?
```{r}
#Group 9 (Unknown) has the longest average stay with 6.18 days.
finaldf %>% group_by(raceethn) %>% summarise(mean(los))
```


16. On average, how much a 20 year-old male white get charged for staying 1 day?
```{r}
#On average he will be charged $3,378
finaldf %>% filter(sex == 1, age == '20', raceethn == '1', los == '1') %>% summarise(mean(trandb))
```


-------

## II. Data Visualization

Continue with the data from part I. 

1. Provides at least 10 meaningful plots. Comments on the plots. All plots should have title, caption, appropriate labels on x and y-axis
```{r}
#This is a chart that graphs the amount teen patients are charged grouped by the sex of the patient
finaldf %>% filter(AgeGroup == 'teen', sex != '9', los < 250) %>% ggplot()+ geom_point(mapping = aes(x = los, y = trandb, color = sex))+ labs(x='Length of Stay', y = 'Total Charged', title = 'Total Room and Board Charged by Sex for Teens', caption = 'Male teens appear to be charged more')
```
```{r}
#This is similar to the previous chart but has patients in the child category to compare.
finaldf %>% filter(AgeGroup == 'child', sex != '9', los < 250) %>% ggplot()+ geom_point(mapping = aes(x = los, y = tot, color = sex))+ labs(x='Length of Stay', y = 'Total Charged', title = 'Total Charged by Sex for Children', caption = 'Not a strong difference seen')
```

```{r}
#This shows the total amount charged broken out by Race/Ethnicity.  This is skewed due to the fact that there are so many more Caucasian patients
finaldf %>% filter(raceethn != '') %>% ggplot()+ geom_col(mapping = aes(x = raceethn, y = trandb))+ labs(x='Race', y = 'Total Charged', title = 'Total Charged grouped by Race/Ethnicity', caption = 'Skewed towards Caucasians')
```

```{r}
#This shows the large disparity in cases for Caucasian
finaldf %>% filter(raceethn != '') %>% ggplot()+ geom_bar(mapping = aes(x = raceethn))+ labs(x='Race', y = 'Count of Patients', title = 'Total Patients Grouped by Race/Ethnicity', caption = 'This should explain the skew towards Caucasians')
```

```{r}
finaldf %>% filter(AgeGroup == 'newborn', sex != '9', b_wt != 0) %>% ggplot()+ geom_point(mapping = aes(x = los, y = b_wt))+ labs(x='Length of Stay', y = 'Body Weight', title = 'Length of Stay vs Newborn body weight', caption = 'Longer stay for babies underweight')
```

```{r}
finaldf %>% filter(er_mode != '0', er_mode != '9', er_mode != '') %>% ggplot()+ geom_bar(mapping = aes(x = er_mode))+ labs(x='Arrival Type', y = 'Count of Arrivals', title = 'Types of Arrivals', caption = 'Ambulance and Personal Arrivals are the Majority')
```


```{r}
finaldf %>% filter(er_mode == '2') %>% ggplot()+ geom_bar(mapping = aes(x = moa))+ labs(x='Month', y = 'Count of Patients', title = 'Helicopter Evacs by Month', caption = 'Late Summer Early Fall have more Helicopter Arrivals')
```

```{r}
finaldf %>% filter(sex != '9', provider == '7214') %>% ggplot()+ geom_bar(mapping = aes(x = sex))+ labs(x='Arrival Type', y = 'Count of Arrivals', title = 'Types of Arrivals', caption = 'Ambulance and Personal Arrivals are the Majority')
```

```{r}
finaldf %>% filter(provider == '7214', sex != '9', age < 60) %>% ggplot()+ geom_histogram(mapping = aes(x = age))+ labs(x='Age', y = 'Count of Patients', title = 'Patient Sex at Women and Infants Hospital', caption = 'Obviously the Male population is signifigantly lower') + facet_wrap(~sex)
```

```{r}
finaldf %>% filter(provider == '7205', sex != '9', age < 60) %>% ggplot()+ geom_histogram(mapping = aes(x = age))+ labs(x='Age', y = 'Count of Patients', title = 'Patient Sex at RI Hospital', caption = 'Pretty equal distrobution') + facet_wrap(~sex)
```


2. Make an animation

```{r}
finaldf %>% filter(provider == '7214', sex != '9', age < 60) %>% ggplot(mapping = aes(x = age))+ geom_histogram()+labs(x='Age', y = 'Count of Patients', title = 'Patient Sex at Women and Infants Hospital', caption = 'Obviously the Male population is signifigantly lower')+transition_states(sex)
```


-------

## III. Predictive Models

Continue with the data from part I. Use the follows as the target and input variables: 
```{r}


#"age","sex","raceethn","provider","moa","mod","admtype","campus", 'los'


finaldf$target <- case_when(
    finaldf$tot < median(finaldf$tot) ~ 'low',
    finaldf$tot >= median(finaldf$tot) ~ 'high',)

preddf <- finaldf

preddf$sex <- factor(preddf$sex)
preddf$raceethn <- factor(preddf$raceethn)
preddf$provider <- factor(preddf$provider)
preddf$moa <- factor(preddf$moa)
preddf$mod <- factor(preddf$mod)
preddf$admtype <- factor(preddf$admtype)
preddf$campus <- factor(preddf$campus)
preddf$target <- factor(preddf$target)

treedf <- preddf %>% select(age, sex, raceethn, provider, moa, mod, admtype,campus, los, target)
names(treedf)[10]<- 'target'

```

*Target Variable*: Create the target variable taking value of 

  - `low` if the total charge of a patient (`tot`) is smaller than the median of the total charge, and

  - `high` otherwise. 
  
```{r}


```


*Input Variables*:

  - "age","sex","raceethn","provider","moa","mod","admtype","campus", 'los'
  
-------

1. Use `filter` function to filter out rows where `raceethn==''` or `admtype==''`. Make sure all the categorical variables are factor, numeric variables are numeric. Set Training : Testing Split = 10 : 90 
```{r}
library(rpart)
library(caret)
treedf <- filter(treedf, raceethn !='', admtype != '')
set.seed(2000)
splitIndex <- createDataPartition(treedf$target, p= .90, list=FALSE)
df_train <- treedf[ splitIndex,]
df_test <- treedf[-splitIndex,]
```

2. Train a decision tree using `rpart`.  Plot the decision tree. Plot the variable importance ranked by the tree. 

```{r}
library(rpart)
library(caret)
library(rattle)
tree_model <- rpart(target ~ ., data = df_train, control = rpart.control(maxdepth = 3))
fancyRpartPlot(tree_model)
tree_model$variable.importance
barplot(tree_model$variable.importance)
pred <- predict(tree_model, df_test, type = 'class')


```
```{r}
conmat <- confusionMatrix(data = pred, reference = df_test$target, positive = 'high')
conmat$overall[1]
```


3. Using caret for this question. Set `Training Control` to be: Use Cross-Validation of 5 folds across all models.  Train & tune at least 3 different models (i.e. three different values for `method=` in the train function of caret).  Plot the hyper-parameter tuning plots for each model. 

```{r}

tuneGrid = expand.grid(maxdepth = 2:10)
trControl = trainControl(method = "cv",
                         number = 10)

model1 <- train(target~., data=df_train, 
                method = "rpart2",
                trControl = trControl,
                tuneGrid = tuneGrid)
#plot(model1)
#pred <- predict(model1, df_test)
#cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "high")
#cm$overall[1]
```

```{r}
#trControl = trainControl(method = "cv",
 #                        number = 10)
#tuneGrid = expand.grid(mtry = 2:4,
 #                      splitrule = c('gini', 'extratrees'),
  #                     min.node.size = c(1:10))

#model2 <- train(target~., data=df_train, 
 #                   method = "ranger", 
  #                  trControl = trControl,
   #                 tuneGrid = tuneGrid)
#plot(model2)

```
```{r}
#plot(model2)
```



4. Plot the comparison of the models in 3. 

5. What is your final selection for the model? Test the accuracy of your final model on the test data. 

6. Create another `target` variable (binary), decide the input variables and redo 1 to 5. 

-------